---
title: "Lab 03 - Nobel laureates"
author: "Cat Seitz"
date: "01.19.23"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1 - examine dataset

Each row is a different Nobel laureate. There are 935 observations with 26 variables. 

### Exercise 2 - create new data frame

Code worked to create a new data frame with 228 observations. 

``` {r make new data frame}

nobel_living <- nobel %>% 
  subset(is.na(died_date)) %>%
  filter(gender != "org") %>%
  drop_na(country)

```

```{r laureate location}

nobel_living <- nobel_living %>%
  mutate(country_us = if_else(country =="USA", "USA", "Other"))

nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```


### Exercise 3

```{r living laureate location}

ggplot(nobel_living_science, aes(x=country_us))+
  geom_bar()+
  facet_wrap(~category)+
  coord_flip()+
  labs(title = "Location of living Nobel laureate winner by field",
       y="Number of winners",
       x="USA vs. other")

```

These results seem to replicate the data shown in the Buzzfeed article (their first graph), but don't necessarily support the overall headline yet. 

### Exercise 4

```{r laureate birth place}

nobel_living_science <- nobel_living_science %>%
  mutate(born_country_us = if_else(born_country =="USA", "USA", "Other"))


ggplot(nobel_living_science, aes(x=born_country_us))+
  geom_bar()

count(nobel_living_science,born_country_us)

```

105 living Nobel laureates were born in the U.s. 

### Exercise 5

```{r laureate location and birth place}

ggplot(nobel_living_science, aes(x=country_us, color=born_country_us, fill=born_country_us))+
  geom_bar()+
  facet_wrap(~category)+
  coord_flip()+
  labs(title = "Location of living Nobel laureate winner by field",
       y="Number of winners",
       x="USA vs. other")

```

Yes, these data support Buzzfeed's claim. While the majority of winners reside in the U.S. when they receive the award, the majority of winners in all disciplines except economics were born outside of the U.S. 

### Exercise 6

```{r}



```

